A Linear Solution of the Linearity Challenge

### 2025-12-26 paper/calculus.tex
LCC

\section{Linear \CP}
\label{sec:calculus}

In this section we give a cursory presentation of \LCP.
%
As we have anticipated in \Cref{sec:introduction}, \LCP is closely related to
\CP~\citep{Wadler14,GayVasconcelos25} except that \LCP features linear channels
instead of sessions. A session consisting of several exchanged messages is
encoded using linear channels by pairing each message with a fresh continuation
channel on which the rest of the session takes place.
%
This way of encoding sessions has been first described by \citet[extended
version]{Kobayashi02b} and later studied by \citet{DardhaGiachinoSangiorgi17}.

We present \LCP starting from its types (\Cref{sec:types}), which coincide with
classical linear logic propositions, then moving on to the syntax of processes
(\Cref{sec:syntax}), their reduction semantics (\Cref{sec:semantics}), the
typing rules (\Cref{sec:typing-rules}) and the formulation of the properties
ensured by the type system (\Cref{sec:properties}).

@@

The interpretation of linear logic propositions as protocols is fairly
usual~\cite{Wadler14}: the multiplicative constants $\Bot$ and $\One$ describe
protocols that terminate with the input/output of a signal; the additive
connectives $A \With B$ and $\A \Plus B$ describe protocols having two possible
continuations of type $A$ and $B$ which are respectively offered/selected; the
multiplicative connectives $A \Par B$ and $A \Ten B$ describe bifurcating
protocols with two continuations of type $A$ and $B$ that are respectively
received and sent. The additive constants $\Zero$ and $\Top$ do not describe
useful protocols, but they can play the role of smallest/largest type in type
systems that support a notion of subtyping~\citep{HornePadovani24}. 
%
The quantifiers $\forall X.A$ and $\exists X.A$ describe the input/output of a
type $X$ with a continuation $A$ and are useful to describe parametric protocol
polymorphism.
%
Finally, the ``of course'' modality $\OfCourse A$ and the ``why not'' modality
$\WhyNot A$ describe shared channels on which servers and clients accept and
request connections of type $A$.

@@

The syntax of processes makes use of an infinite set of \emph{channels}, ranged
over by $x$, $y$ and $z$, and is shown in~\Cref{tab:syntax}.
%
A link $\Link\x\y$ denotes the merging of the channels $x$ and $y$, so that the
message sent on one of the channels is forwarded to the other. As discussed in
the literature~\cite{LindleyMorris15}, this form is useful for modeling
higher-order sessions, that is the exchange of an existing channel on another
channel.
%
The processes $\Wait\x.P$ and $\Close\x$ respectively model the input and output
of a termination signal on the channel $x$. The latter process has no
continuation, while the former process continues as $P$ once the signal has been
received.
%
The process $\Fail\x$ can be used to denote a failure concerning the channel
$x$.
%
The processes $\Case\x\z{Q_1}{Q_2}$ and $\Select\x{\InTag_i}\z.P$ for $i=1,2$
describe the input/output of a label $\InTag_i$ along with a fresh continuation
$z$ on the channel $x$. The receiver continues as $Q_i$ depending on the label
$\InTag_i$. In practice it is useful to consider arbitrary (domain-specific)
sets of labels instead of the canonical $\InTag_1$ and $\InTag_2$. This
generalisation has no substantial impact on the formalisation, so we stick to
binary choices in line with the usual presentations of linear logic.
%
The processes $\Join\x\y\z.R$ and $\Fork\x\y\z{P}{Q}$ describe the input/output
of a pair of fresh channels $y$ and $z$ on the channel $x$. The former process
has a single continuation $R$ that can use both $y$ and $z$. The latter process
forks into $P$ and $Q$, each using $y$ and $z$ respectively. These forms are
often used in conjunction with links to model higher-order sessions.
%
The processes $\Forall\x\X\z.P$ and $\Exists\x{A}\z.P$ describe the input/output
of a type on the channel $x$ along with a fresh continuation $z$.

Next we have process forms dealing with shared (non-linear) channels. The
$\Server\x\y.P$ and $\Client\x\y.P$ processes respectively denote \emph{servers}
and \emph{clients} acting on the shared channel $x$. Each request (from a
client) spawns a copy of the server's body using the continuation channel $y$.
The process $\Weaken\x.P$ denotes an explicit \emph{weakening}, that is a client
that \emph{does not} use $x$. The process $\Contract\x\y\z.P$ denotes an
explicit \emph{contraction} whereby a client uses $x$ multiple times (once with
name $y$ and once with name $z$).

@@

\begin{example}
    \label{ex:booleans}
    We can represent the type of a boolean value being produced as the protocol
    $\Bool \eqdef \One \Plus \One$ and that of a boolean value being consumed as
    its dual $\dual\Bool = \Bot \With \Bot$. Following these types, the boolean
    constants can be modeled by the processes
    \[
        \TrueP(x) \eqdef \Select\x\LeftTag{x'}.\Close{x'}
        \qquad
        \FalseP(x) \eqdef \Select\x\RightTag{x'}.\Close{x'}
    \]
    and the boolean negation function by the process
    \[
        \NotP(x,y) \eqdef
        \BigCase\x{x'}{
            \Wait{x'}.\FalseP(y)
        }{
            \Wait{x'}.\TrueP(y)
        }
    \]

    As an example, the composition $\Cut[\Bool]\x{\TrueP(x)}{\NotP(x,y)}$
    produces false on $y$.
    %
    Note the use of explicit continuations in these processes and the fact that
    each channel is used exactly once. The same processes in \CP would be
    written as
    \[
        \TrueP(x) \eqdef \Select\x\LeftTag{}.\Close\x
        \quad
        \FalseP(x) \eqdef \Select\x\RightTag{}.\Close\x
        \quad
        \NotP(x,y) \eqdef
        \BigCase\x{}{
            \Wait\x.\FalseP(y)
        }{
            \Wait\x.\TrueP(y)
        }
    \]
    where each channel is used multiple times to indicate the sequence of
    input/output actions pertaining to the same session.
    %
    In general, the \CP version of an \LCP process can be obtained by reusing
    the same channel $x$ in place of the continuation $z$ in \Cref{tab:syntax}.
    %
    \eoe
\end{example}

@@

The operational semantics of \Calculus is shown in~\Cref{tab:semantics} and
is given by two relations: a \emph{structural pre-congruence} relation $\pcong$
and a \emph{reduction} relation $\red$. Structural pre-congruence relates
processes that we consider to be essentially indistinguishable. Reduction
relates processes in which a communication occurs. In some presentations of
linear logic~\cite{BaeldeDoumaneSaurin16,Doumane17}, these relations are
respectively dubbed \emph{external} and \emph{internal} reductions. Let us
describe the two relations more in detail.

@@

There is no pre-congruence rule asserting that parallel composition is
associative. While this is a sound property of parallel composition, it turns
out to be unnecessary for proving the properties of \LCP that we consider. We
leave it out in light of the forthcoming formalisation of the calculus, in which
every simplification is welcome, although standard presentations of classical
linear logic admit such law~\cite{Wadler14,Doumane17}.
%
Another rule that deserves attention is \SServer. In this case, a server prefix
$\Server\y\u$ can be pulled out of a cut only provided that the other process in
the cut is also a server on the channel restricted by the cut.

Reduction is defined by the \rulename{r-*} rules.
%
The reduction rules \RLink, \RClose, \RSelect, \RFork, \RExists, \RConnect,
\RWeaken and \RContract are in one-to-one correspondence with the well-known
reductions of session calculi and the principal cut reductions of linear logic.
\RLink eliminates a link $\Link\x\y$ by substituting $y$ for $x$. \RClose closes
a session $x$ when a termination signal is exchanged on $x$. \RSelect and \RFork
respectively model the communication of a tag $\InTag_i$ and of a pair of fresh
continuation channels. \RExists models the instantiation of a polymorphic
variable $X$ as the communication of a type $A$. We write $Q\subst{A}\X$ for the
process obtained by replacing every free occurrence of the type variable $X$
with $A$. Each of these communications consumes the channel $x$ and possibly
creates and transfers \emph{fresh continuations} on which the rest of the
conversation takes place, in line with the encoding of sessions into the linear
$\pi$-calculus~\cite{KobayashiPierceTurner99,Kobayashi02b,DardhaGiachinoSangiorgi17}.
%
The rules \RConnect, \RWeaken and \RContract describe interactions on shared
channels and respectively represent the connection between a server and a
client, the disposal of an unused server, and the duplication of a server. In
\RWeaken and \RContract we use some convenient (albeit slightly informal)
notation for denoting sequences of (pairwise distinct) channels and prefixes. In
particular, $\seqof\z$ stands for a sequence $z_1, \dots, z_n$ of channels,
$\Weaken{\seqof\z}.Q$ stands for $\Weaken{z_1}\dots\Weaken{z_n}.Q$ and
$\Contract{\seqof\z}{\seqof\z'}{\seqof\z''}.R$ stands for
$\Contract{z_1}{z_1'}{z_1''}\dots\Contract{z_n}{z_n'}{z_n''}.R$.

@@

>>>>

\section{A Linear Calculus of Continuations}
\label{sec:calculus}

In this section we give a cursory presentation of \Calculus starting from its
types (\Cref{sec:types}), which coincide with classical linear logic
propositions, then moving on to the syntax of processes (\Cref{sec:syntax}),
their reduction semantics (\Cref{sec:semantics}), the typing rules
(\Cref{sec:typing-rules}) and the formulation of the properties ensured by the
type system (\Cref{sec:properties}).

As we have anticipated in \Cref{sec:introduction}, \Calculus is closely related
to \CP~\citep{Wadler14,GayVasconcelos25} except that \Calculus features linear
channels instead of sessions.
%
We refer the reader to the literature on
\CP~\cite{Wadler14,LindleyMorris15,GayVasconcelos25} and other session calculi
based on linear logic~\cite{CairesPfenning10,CairesPfenningToninho16} for a
thorough introduction to these models.

@@

The interpretation of linear logic propositions as behaviours is fairly
standard. Constants and connectives describe \emph{linear channels}, which must
be used for a single communication, whereas the modalities $\OfCourse$ and
$\WhyNot$ describe \emph{shared channels}.
%
The multiplicative constants $\Bot$ and $\One$ describe channels used for
receiving/sending an empty message (without continuations).
%
The additive connectives $A \With B$ and $A \Plus B$ describe  channels used for
receiving/sending either a continuation of type $A$ or a continuation of type
$B$. The sender selects one of the two possibilities, while the receiver offers
both.
%
The multiplicative connectives $A \Par B$ and $A \Ten B$ describe channels used
for receiving/sending two continuations, one of type $A$ and the other of type
$B$.
%
The additive constants can play the role of smallest/largest type in type
systems that support a notion of subtyping~\citep{HornePadovani24}. We will not
be using them in this work.
%
The quantifiers $\forall X.A$ and $\exists X.A$ describe channels used for
receiving/sending a type $X$ along with a continuation of type $A$. They are
useful to describe parametric protocol polymorphism.
%
Finally, the ``of course'' modality $\OfCourse A$ and the ``why not'' modality
$\WhyNot A$ describe shared channels on which servers and clients accept and
request connections of type $A$.

@@

The syntax of processes makes use of an infinite set of \emph{channels}, ranged
over by $x$, $y$ and $z$, and is shown in~\Cref{tab:syntax}.
%
A link $\Link\x\y$ denotes the merging of the channels $x$ and $y$, so that each
message sent on one of the channels is forwarded to the other. As discussed in
the literature~\cite{LindleyMorris15}, this form is useful for modeling
higher-order sessions, that is the exchange of an existing channel on another
channel.
%
The processes $\Wait\x.P$ and $\Close\x$ respectively model the input and output
of an empty message on the channel $x$. The latter process has no continuation,
while the former process continues as $P$ once the message has been received.
%
The process $\Fail\x$ can be used to denote a failure concerning the channel
$x$.
%
The process $\Case\x\z{Q_1}{Q_2}$ offers a choice on channel $x$ and continues
as either $Q_1$ or $Q_2$ with $z$ bound to the received continuation depending
on which branch is selected.
%
The process $\Select\x{\InTag_i}\z.P$ performs a choice (represented by a label
$\InTag_i$ with $i=1,2$) and sends a fresh continuation $z$ on the channel $x$.
%
The processes $\Join\x\y\z.R$ and $\Fork\x\y\z{P}{Q}$ describe the input/output
of two fresh continuations channels $y$ and $z$ on the channel $x$. The receiver
can use $y$ and $z$ in whatever order. The sender forks into $P$ and $Q$, each
using $y$ and $z$ respectively. These forms are often used in conjunction with
links to model higher-order sessions.
%
The processes $\Forall\x\X\z.P$ and $\Exists\x{A}\z.P$ describe the input/output
of a type on the channel $x$ along with a fresh continuation $z$.

Next we have process forms dealing with shared (non-linear) channels. The
processes $\Server\x\y.P$ and $\Client\x\y.P$ respectively denote \emph{servers}
and \emph{clients} acting on the shared channel $x$. Each request (from a
client) spawns a copy of the server's body using the continuation channel $y$.
The process $\Weaken\x.P$ denotes an explicit \emph{weakening}, that is a client
that \emph{does not} use $x$. The process $\Contract\x\y\z.P$ denotes an
explicit \emph{contraction} whereby a client uses $x$ multiple times (once with
name $y$ and once with name $z$).

@@

\begin{example}
    \label{ex:booleans}
    We can represent the protocol of a boolean value being produced as the type
    $\Bool \eqdef \One \Plus \One$ and that of a boolean value being consumed as
    its dual $\dual\Bool = \Bot \With \Bot$. Following these types, the boolean
    constants can be modeled by the processes
    \[
        \TrueP(x) \eqdef \Select\x\LeftTag{x'}.\Close{x'}
        \qquad
        \FalseP(x) \eqdef \Select\x\RightTag{x'}.\Close{x'}
    \]
    and the boolean negation function by the process
    \[
        \NotP(x,y) \eqdef
        \BigCase\x{x'}{
            \Wait{x'}.\FalseP(y)
        }{
            \Wait{x'}.\TrueP(y)
        }
    \]

    As an example, the composition $\Cut[\Bool]\x{\TrueP(x)}{\NotP(x,y)}$
    produces false on $y$.
    %
    Note the use of explicit continuations in these processes and the fact that
    each channel is used exactly once. The same processes in \CP would be
    written as
    \[
        \TrueP(x) \eqdef \Select\x\LeftTag{}.\Close\x
        \quad
        \FalseP(x) \eqdef \Select\x\RightTag{}.\Close\x
        \quad
        \NotP(x,y) \eqdef
        \BigCase\x{}{
            \Wait\x.\FalseP(y)
        }{
            \Wait\x.\TrueP(y)
        }
    \]
    where each channel is used multiple times to indicate the sequence of
    input/output actions pertaining to the same session.
    %
    In general, the \CP version of an \Calculus process can be obtained by reusing
    the same channel $x$ in place of the continuation $z$ in \Cref{tab:syntax}.
    %
    \eoe
\end{example}

@@

The operational semantics of \Calculus is shown in~\Cref{tab:semantics} and is
given by two relations: a \emph{structural pre-congruence} relation $\pcong$ and
a \emph{reduction} relation $\red$. Structural pre-congruence relates processes
that we consider to be essentially indistinguishable. Reduction relates
processes in which a communication occurs. Let us describe the two relations
more in detail.

@@

There is no pre-congruence rule asserting that parallel composition is
associative. While this is a sound property of parallel composition, it turns
out to be unnecessary for proving the properties of \Calculus that we consider. We
leave it out in light of the forthcoming formalisation of the calculus, in which
every simplification is welcome, although standard presentations of classical
linear logic admit such law~\cite{Wadler14,Doumane17}.
%
Another rule that deserves attention is \SServer. In this case, a server prefix
$\Server\y\u$ can be pulled out of a cut only provided that the other process in
the cut is also a server on the channel restricted by the cut.

Reduction is defined by the \rulename{r-*} rules.
%
The reduction rules \RLink, \RClose, \RSelect, \RFork, \RExists and \RConnect
are in one-to-one correspondence with the well-known reductions of session
calculi and the principal cut reductions of linear logic. These reductions erase
the topmost cut and replace it with zero, one or two new cuts, depending on the
number of continuations that are exchanged. Note that \RLink eliminates a link
$\Link\x\y$ by substituting $y$ for $x$ in the scope of $x$ and \RClose models
the communication of an empty message (without continuations), hence no new cut
is present in the reduct. \RSelect, \RFork and \RConnect are standard and
\RExists models the instantiation of a polymorphic variable $X$ as the
communication of a type $A$. We write $Q\subst{A}\X$ for the process obtained by
replacing every free occurrence of the type variable $X$ with $A$.
%
The reduction rules \RWeaken and \RContract respectively represent the disposal
of an unused server and the duplication of a server. In these rules we use some
convenient (albeit slightly informal) notation for denoting sequences of
(pairwise distinct) channels and prefixes. In particular, $\seqof\z$ stands for
a sequence $z_1, \dots, z_n$ of channels, $\Weaken{\seqof\z}.Q$ stands for
$\Weaken{z_1}\dots\Weaken{z_n}.Q$ and
$\Contract{\seqof\z}{\seqof\z'}{\seqof\z''}.R$ stands for
$\Contract{z_1}{z_1'}{z_1''}\dots\Contract{z_n}{z_n'}{z_n''}.R$.

@@

\newcommand\MonoEchoS{\mathit{Server}}
\newcommand\MonoEchoC{\mathit{Client}}
\begin{example}
    \label{ex:neg}
    %
    We illustrate an example of higher-order session by modeling an echo server
    that consumes boolean values and sends them back unchanged. The protocol
    implemented by the server is described by the type $\OfCourse(\dual\Bool
    \Par (\Bool \Ten \One))$ where the modality~$\OfCourse$ indicates that the
    server is able to accept an unbounded number of requests and the type
    $\dual\Bool \Par (\Bool \Ten \One)$ describes the sequence of actions
    performed by the server at each connection with a client: the server first
    consumes a boolean value (say $u$, of type $\dual\Bool$), then produces
    another boolean value (say $v$, of type $\Bool$) and finally sends an empty
    message. We can model the server and a possible client thus:
    \begin{align*}
        \MonoEchoS(x) & \eqdef
        \Server\x\y.
        \Join\y\u{y'}.
        \Fork{y'}\v{y''}{
            \Link\v\u
        }{
            \Close{y''}
        } \\
        \MonoEchoC(x,z) & \eqdef
        \Client\x\y.
        \Fork\y\u{y'}{
            \TrueP(u)
        }{
            \Join{y'}\v{y''}.
            \Wait{y''}.
            \Link\v\z
        }
    \end{align*}

    Notice the use of continuations for chaining communications together. In the
    server, the link $\Link\v\u$ merges $v$ and $u$ so as to send the same
    channel $u$ received from the client. In the client, the link $\Link\v\z$
    ``assigns'' the message $v$ received from the server to the free channel
    $z$, which represents the result of the interaction.
    %
    If we write $\reds$ for the reflexive, transitive closure of $\red$ it is
    easy to verify that $\Cut\x{\MonoEchoC(x)}{\MonoEchoS(x)} \red^* \TrueP(z)$,
    as expected.
    %
    \eoe
\end{example}

### 2025-12-26 paper/introduction.tex
LCC

One of the challenges in the benchmark, henceforth called \emph{linearity
challenge}, concerns the formalisation of a \emph{minimal calculus of sessions}.
%
Sessions and session types~\citep{Honda93,HondaVasconcelosKubo98,HuttelEtAl16}
are established abstractions for the static analysis of distributed programs
based on peer-to-peer communication channels. Every session type system revolves
around three key ideas: (1) session endpoints are \emph{linear resources} that
cannot be discarded or duplicated without compromising some safety and liveness
properties of a program; (2) the type of a session endpoint is \emph{updated}
after each use to reflect the state of the protocol it describes; (3) peer
session endpoints are meant to be used in complementary ways so as to guarantee
the absence of communication errors and, to some extent, progress of the
interaction.
%
The linearity challenge is motivated by the observation that the proper
management of linear resources in a formalisation often requires a large number
of auxiliary definitions and technical results that divert from the main problem
under study~\cite{CarboneEtAl24}. One of the alleged culprits of such complexity
is \emph{context splitting}, namely the operation that partitions a typing
context in such a way that the linear resources described therein end up in only
one of the partitions. This observation has led to the exploration of various
alternative techniques including leftover typing~\cite{ZalakainDardha21}, the
use of linearity predicates~\cite{SanoKavanaghPientka23} and tagged
contexts~\cite{ZackonEtAl25}.

In this paper we approach the linearity challenge from a different angle:
\emph{instead of proposing new techniques that make it easy to formalise the
calculus in the challenge, we propose a (relatively) new calculus that is easy
to formalise with the existing techniques}.
%
More specifically, we describe a calculus called \LCP whose type system
coincides with the proof system of classical linear logic and that features
\emph{linear channels} instead of sessions. Linear channels are meant to be used
for a \emph{single} communication. Structured communications involving the
exchange of several messages within the same session are modeled in \LCP using
\emph{explicit continuations} according to the well-known encoding of (binary)
sessions into the linear
$\pi$-calculus~\citep{KobayashiPierceTurner99,Kobayashi02b,DardhaGiachinoSangiorgi17}.
%
Overall, \LCP is a linear version of \CP~\cite{Wadler14} in which sessions are
encoded instead of being featured natively.

The logical foundations of \LCP and the use of explicit continuations play an
important role in taming the complexity of the formalisation.
%
Using a calculus based on linear logic prevents \emph{by construction} the same
(sequential) process to own both endpoints of a session. This situation is
clearly undesirable since it does not correspond to a useful pattern of
interaction (every meaningful session requires its endpoints to be used by
parallel processes) and is a potential source of deadlocks.
%
From the standpoint of the formalisation, where the representation of channels
is a primary design choice, it spares us the need to distinguish the two
endpoints of a session, \eg by means of
polarities~\citep{GayHole05,ZalakainDardha21} or by using different names
connected by the same binder~\citep{Vasconcelos12,GayVasconcelos25}.
%
Using a calculus with linear channels and explicit continuations spares us the
need to \emph{update} the type of channels in typing contexts. Once a channel
has been used it is effectively consumed, therefore its type can be
\emph{removed} from the typing context and the type of the continuation channel
is \emph{added} back to the typing context. As is turns out, removing and adding
types is easier than updating them. Even more so considering that the type of
continuation channels must be added \emph{at the beginning} of a typing context,
since such channels are fresh by definition.

>>>>

One of the challenges in the benchmark, henceforth called \emph{linearity
challenge}, concerns the formalisation of a \emph{minimal calculus of sessions}.
%
Sessions and session
types~\citep{Honda93,HondaVasconcelosKubo98,HuttelEtAl16,GayVasconcelos25} are
established abstractions for the static analysis of distributed programs based
on peer-to-peer communications. Every session type system revolves around three
key ideas: (1) session endpoints are \emph{linear resources} that cannot be
discarded or duplicated without compromising some safety and liveness properties
of a program; (2) the type of a session endpoint is \emph{updated} after each
use to reflect the state of the protocol it describes; (3) peer session
endpoints are meant to be used in complementary ways so as to guarantee the
absence of communication errors and, to some extent, progress of the
interaction.
%
The linearity challenge is motivated by the observation that the proper
management of linear resources in a formalisation often requires a large number
of auxiliary definitions and technical results that divert from the main problem
under study~\cite{CarboneEtAl24}. One of the alleged culprits of such complexity
is \emph{context splitting}, namely the operation that partitions a typing
context in such a way that the linear resources described therein end up in only
one of the partitions. This observation has led to the exploration of various
alternative techniques including leftover typing~\cite{ZalakainDardha21}, the
use of linearity predicates~\cite{SanoKavanaghPientka23} and tagged
contexts~\cite{ZackonEtAl25}.

In this paper we approach the linearity challenge from a different angle:
\emph{instead of proposing new techniques that make it easy to formalise the
calculus in the challenge, we propose a (relatively) new calculus that is easy
to formalise with the existing techniques}.
%
More specifically, we describe the Linear Calculus of Continuations (\LCC) whose
type system coincides with the proof system of classical linear logic and that
features \emph{linear channels} instead of sessions. Linear channels are meant
to be used for a \emph{single} communication. Structured conversations can be
modeled in \LCC using \emph{explicit continuations} as in the well-known
encoding of (binary) sessions into the linear
$\pi$-calculus~\citep{KobayashiPierceTurner99,Kobayashi02b,DardhaGiachinoSangiorgi17}:
each message exchanged on a linear channel may include one or two fresh channels
-- the \emph{continuations} -- on which the rest of the conversation takes
place.
%
Overall, \LCC is nothing but a lower-level, linear version of
\CP~\cite{Wadler14}, Wadler's calculus of sessions based on classical linear
logic, such that sessions are encoded instead of being featured natively.

The logical foundations of \LCC and the use of explicit continuations play an
important role in taming the complexity of the formalisation.
%
Using a calculus based on linear logic prevents \emph{by construction} the same
(sequential) process to own both endpoints of a session. This situation is
clearly undesirable since it does not correspond to a useful pattern of
interaction (every meaningful session requires its endpoints to be used by
parallel processes) and is a potential source of deadlocks.
%
From the standpoint of the formalisation, where the representation of channels
is a primary design choice, it spares us the need to distinguish the two
endpoints of a session, \eg by means of
polarities~\citep{GayHole05,ZalakainDardha21} or by using different names
connected by the same binder~\citep{Vasconcelos12,GayVasconcelos25}.
%
Using a calculus with linear channels and explicit continuations spares us the
need to \emph{update} the type of channels in typing contexts. Once a channel
has been used it is effectively consumed, therefore its type can be
\emph{removed} from the typing context and the type of the continuation channel
is \emph{added} back to the typing context. As is turns out, removing and adding
types is easier than updating them. Even more so considering that the type of
continuation channels must be added \emph{at the beginning} of a typing context,
since such channels are fresh by definition.
